\thesistitle

### 2019-08-16 sections/background.tex
Sander

\subsection{Type 4 clones}
For this study we have chosen to take type 4 clones out of the scope, because they are both hard to detect and hard to refactor. A study by Kodhai et al \cite{kodhai2013method} looks into the distribution of the different types of clones in several open source systems (see table 6 of his study). It becomes apparent that type 4 clones exist way less in source code than all of the other types of clones. For instance, for the J2sdk-swing system he finds 8115 type 1 clones, 8205 type 2 clones, 11209 type 3 clones and only 30 type 4 clones. Because of that, we can conclude that type 4 clones are relatively less relevant to study.

@@

\subsection{Clone refactoring in relationship to its context}
How to refactor clones is highly dependent on their context. Method-level clones can be extracted to a method \cite{kodhai2013method} if all occurrences of the clone reside in the same class. If a method level clone is duplicated among classes in the same inheritance structure, we might need to pull-up a method in the inheritance structure. If instances of a method level clone are not in the same inheritance structure, we might need to either make a static method or create an inheritance structure ourselves. So not only a single instance of a clone has a context, but also the relationship between individual instances in a clone class. This is highly relevant to the way in which the clone has to be refactored.

@@

\section{(Object-Oriented) Programming Languages}
In this section we describe the relevant parts of the programming languages we research and experiment on.

>>>>

\subsection{Type 4 clones}
For this study we have chosen to take type 4 clones out of the scope, because they are both hard to detect and hard to refactor. % TODO: Sander - why is this the case?
A study by Kodhai et al \cite{kodhai2013method} looks into the distribution of the different types of clones in several open source systems (see table 6 of his study). It becomes apparent that type 4 clones exist way less in source code than all of the other types of clones. For instance, for the J2sdk-swing system he finds 8115 type 1 clones, 8205 type 2 clones, 11209 type 3 clones and only 30 type 4 clones. Because of that, we can conclude that type 4 clones are relatively less relevant to study.
% TODO: Sander - Deze subsectie wil ik wel even over hebben, loopt niet helemaal lekker maar niet zo makkelijk uit te leggen

@@

\subsection{Clone refactoring in relationship to its context}
% TODO: Sander - Deze subsectie lijkt hier niet helemaal op zn plek
How to refactor clones is highly dependent on their context. Method-level clones can be extracted to a method \cite{kodhai2013method} if all occurrences of the clone reside in the same class. If a method level clone is duplicated among classes in the same inheritance structure, we might need to pull-up a method in the inheritance structure. If instances of a method level clone are not in the same inheritance structure, we might need to either make a static method or create an inheritance structure ourselves. So not only a single instance of a clone has a context, but also the relationship between individual instances in a clone class. This is highly relevant to the way in which the clone has to be refactored.

@@

\section{(Object-Oriented) Programming Languages}
% TODO: Sander - leg ook nog even in 1 zin uit wat een lezer aan al deze info heeft
In this section we describe the relevant parts of the programming languages we research and experiment on.

### 2019-08-16 sections/introduction.tex
Sander

\chapter{Introduction}
%\todo[inline,color=blue!10]{Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
%Problem Analysis: Here you present your analysis of the problem situation that your research will address.
%How does this problem manifest itself at your host organisation?
%Also summarises existing scientific insight into the problem.}
\label{ch:introduction}
Refactoring is used to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler2018refactoring, wake2004refactoring}. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. However, there are also refactoring opportunities that are rather trivial and repetitive to execute. In this thesis, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve maintainability of the refactored code.

Duplication in source code is often seen as one of the most harmful types of technical debt. \textit{``Clones are problematic for the maintainability of a program, because if the clone is altered at one location to correct an erroneous behaviour, you cannot be sure that this correction is applied to all the cloned code as well. Additionally, the code base size increases unnecessarily and so increases the amount of code to be handled when conducting maintenance work''} \cite{ostberg2014automatically}. Bruntink et al.~\cite{bruntink2005use} show that code clones can contribute up to 25\% of the code size.

@@

\section{Problem statement}
In this section we describe the problem we address with this study and the research questions that we answer in order to contribute to solving the problem.

@@

The process of improving maintainability through the refactoring of duplicate code is time consuming and error-prone. This process mainly consists of these two aspects:
\begin{itemize}
	\item Find refactoring candidates, either tool-assisted or manually.
	\item Refactor identified candidates, either tool-assisted or manually.
\end{itemize}
There are tools that assist in the process of finding duplicate code, but none of these tools identify all duplication problems. Often, these clone detection tools result in false-positives and false negatives \cite{roy2007survey}. We define false negatives as clones that can and should be refactored but are not found due to the configuration by which they are detected. We define false-positives as clones that do not improve the system maintainability when refactored. Many false negatives are due to the thresholds that are used. Many clone detection tools \cite{sajnani2016sourcerercc, svajlenko2016bigcloneeval} only consider clones that have a minimum of 6 lines. Using such an approach is bound to result in many false negatives, but using a lower threshold is shown to increase the number of false-positives. %TODO cite threshold paper

A study by Batova et al. \cite{bavota2012does} shows that the process of refactoring often leaves side effects in the code. This study reports that refactorings involving hierarchies (e.g., pull up method), tend to induce faults very frequently. \textit{This suggests more accurate code inspection or testing activities when such specific refactorings are performed.} Such specific refactorings often have to be used when dealing with duplication in software \cite{fowler2018refactoring, fontana2015duplicated}. Because of that, refactoring code clones has been empirically proven to cause bugs or other side effects in code.

@@

\section{Contributions}
Many studies report that code clones negatively affect maintainability \cite{heitlager2007practical, monden2002software, juergens2009code, chatterji2013effects}. However, no studies yet show in what cases code clones can reduce maintainability in source code. Refactoring often includes tradeoffs between design alternatives. With some code clones, the refactored alternative is worse maintainable than keeping the duplication \cite{kapser2006cloning, aversano2007clones, hotta2010duplicate, kim2005empirical, krinke2007study, saha2010evaluating}. In this study, we analyze the maintainability of refactored code clones, in order to improve the suggestion of code clones that should be refactored. This assists with both the identification and refactoring of code clones.

\subsection{Identification}
There are many tools to detect clones. The goal of most of these tools is to assist developers in reducing duplication in their code, i.e. assisting in the refactoring process. The problem is that these tools have no limited insight on the impact of refactoring such clones on the design of the software. In this study, we can analyze a before- and after-refactoring snapshot of the code to determine the impact. A higher maintainability after refactoring increases the support for the clone being a true-positive. This way the results of our study can support the clone identification process.

@@

\section{Scope}
In this study, we perform research efforts to be able to detect refactorable clones. We will apply refactoring techniques to a subset of these clones and analyze the maintainability of the resulting source code.

@@

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. In Chapter~\ref{chap:clonetypes} we list shortcomings with clone definitions from literature and propose a set of clone type definitions which can be automatically refactored. In Chapter \ref{ch:clonerefactor} we propose a tool to detect, analyze and automatically refactor such clones that can be refactored. Using this tool we perform a set of experiments, of which the results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work} contains the work related to this thesis. Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

>>>>

\chapter{Introduction}
%\todo[inline,color=blue!10]{Context: what is the bigger scope of the problem you are trying to solve? Try to connect to societal/economical challenges.
%Problem Analysis: Here you present your analysis of the problem situation that your research will address.
%How does this problem manifest itself at your host organisation?
%Also summarises existing scientific insight into the problem.}
\label{ch:introduction}
% TODO: Sander - What is refactoring? Hier zeg je meteen waarvoor het gebruikt wordt
Refactoring is used to improve quality related attributes of a codebase (maintainability, performance, etc.) without changing the functionality. There are many methods that have been introduced to help with the process of refactoring \cite{fowler2018refactoring, wake2004refactoring}. However, most of these methods still require manual assessment of where and when to apply them. Because of this, refactoring takes up a signification portion of the development process \cite{lientz1978characteristics, mens2004survey}, or does not happen at all \cite{mens2003refactoring}. For a large part, refactoring requires domain knowledge to do it right. 
However, there are also refactoring opportunities that are rather trivial and repetitive to execute. 
% De zien hierna staat ook al in laatste alineas, past minder goed binnen intro imo
In this thesis, we take a look at the challenges and opportunities in automatically refactoring duplicated code, also known as ``code clones''. The main goal is to improve maintainability of the refactored code.

Duplication in source code is often seen as one of the most harmful types of technical debt. \textit{``Clones are problematic for the maintainability of a program, because if the clone is altered at one location to correct an erroneous behaviour, you cannot be sure that this correction is applied to all the cloned code as well. Additionally, the code base size increases unnecessarily and so increases the amount of code to be handled when conducting maintenance work''} \cite{ostberg2014automatically}. Bruntink et al.~\cite{bruntink2005use} show that code clones can contribute up to 25\% of the code size. % TODO: Sander - why is it bad to increase code size (i know ;) but reader might not)

@@

\section{Problem statement}
In this section we describe the problem we address in this study and the research questions that we answer in order to contribute to solving the problem.

@@

The process of improving maintainability through the refactoring of duplicate code is time consuming and error-prone. This process mainly consists of these two aspects:
\begin{itemize}
	\item Find refactoring candidates, either tool-assisted or manually. % TODO: Sander - is tool assisted niet ook beetje manual? lijkt meer tool-assisted vs fully manual of iets
	\item Refactor identified candidates, either tool-assisted or manually.
\end{itemize}
There are tools that assist in the process of finding duplicate code, but none of these tools identify all duplication problems. Often, these clone detection tools result in false-positives and false negatives \cite{roy2007survey}. We define false negatives as clones that can and should be refactored but are not found due to the configuration by which they are detected. We define false-positives as clones that do not improve the system maintainability when refactored. Many false negatives are due to the thresholds that are used. Many clone detection tools \cite{sajnani2016sourcerercc, svajlenko2016bigcloneeval} only consider clones that have a minimum of 6 lines. Using such an approach is bound to result in many false negatives, but using a lower threshold is shown to increase the number of false-positives. %TODO cite threshold paper

A study by Batova et al. \cite{bavota2012does} shows that the process of refactoring often leaves side effects in the code. This study reports that refactorings involving hierarchies (e.g., pull up method) % TODO: Sander -unsure of pull up method bekende term is, miss "creating a new method in a superclass"
, tend to induce faults very frequently. \textit{This suggests more accurate code inspection or testing activities % TODO: Sander - .. are needed? 
when such specific refactorings are performed.} Such specific refactorings often have to be used when dealing with duplication in software \cite{fowler2018refactoring, fontana2015duplicated}. Because of that, refactoring code clones has been empirically proven to cause bugs or other side effects in code.

@@

\section{Contributions}
Many studies report that code clones negatively affect maintainability \cite{heitlager2007practical, monden2002software, juergens2009code, chatterji2013effects}. However, no studies yet show in what cases code clones can reduce maintainability in source code. Refactoring often includes tradeoffs between design alternatives. With some code clones, the refactored alternative is less maintainable than keeping the duplication \cite{kapser2006cloning, aversano2007clones, hotta2010duplicate, kim2005empirical, krinke2007study, saha2010evaluating}. In this study, we analyze the maintainability of refactored code clones, in order to improve the suggestion of code clones that should be refactored. This assists with both the identification and refactoring of code clones.

\subsection{Identification}
There are many tools to detect clones. The goal of most of these tools is to assist developers in reducing duplication in their code, i.e. assisting in the refactoring process. The problem is that these tools have no limited % TODO: Sander -of limited of no, niet beide ;)
insight on the impact of refactoring such clones on the design of the software. In this study, we can analyze a before- and after-refactoring snapshot of the code to determine the impact. A higher maintainability after refactoring increases the support for the clone being a true-positive. This way the results of our study can support the clone identification process.

@@

\section{Scope}
% TODO: Sander - iets over OOP hier?
In this study, we perform research efforts to be able to detect refactorable clones. We will apply refactoring techniques to a subset of these clones and analyze the maintainability of the resulting source code.

@@

\section{Outline}
In Chapter~\ref{ch:background} we describe the background of this thesis. In Chapter~\ref{chap:clonetypes} we list shortcomings with clone definitions from literature and propose a set of clone type definitions which can be automatically refactored. % TODO: Sander - Nu zeg je dat de definities automatisch gerefactored kunnen worden
In Chapter \ref{ch:clonerefactor} we propose a tool to detect, analyze and automatically refactor such clones that can be refactored. Using this tool we perform a set of experiments, of which the results are shown in Chapter~\ref{ch:results} and discussed in Chapter~\ref{ch:discussion}. Chapter~\ref{ch:related_work} contains the work related to this thesis. Finally, we present our concluding remarks in Chapter~\ref{ch:conclusion} together with future work.

### 2019-08-16 sections/research.tex
Sander

\chapter{Defining refactoring-oriented clone types}\label{chap:clonetypes}
In Section~\ref{chap:backgroundclonetypes} we introduced the four clone types as defined in literature. These simple definitions are suitable for analysis of a codebase. Their detection results in simple to understand numbers to argue about a codebase. However, these clone types have a few flaws which makes it hard to argue to what extend two fragments of code can and should be refactored. For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings to increase the chance that we can refactor the clone while improving the design.

We also look into clone detection tools for their suitability to support the proposed clone type definitions. We selected a few criteria  Most clone detection tools support these definitions of clone types. However, many of these tools use a vastly different approach. A study by Saini et al \cite{saini2018towards} outlines different clone detection tools and compares their results for each of type 1-3 clones. Even though they operate on the same type definitions, the tools used in this study yield different results.

\section{Shortcomings of clone types}
Clone types 1-3 (further explained in Section~\ref{chap:backgroundclonetypes}) allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. This has allowed for large scale analyses of duplication~\cite{livieri2007very}. However, these clone type definitions have shortcomings which makes the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

\section{Refactoring-oriented clone types}\label{sec:rtypes}
To resolve the shortcomings of clone types as outlined in the previous section, we propose alternative definitions for clone types directed at detecting clones that can and should be refactored. We have named these clones T1R (type 1R), T2R and T3R clones. These definitions address problems of the corresponding literature definitions. The ``R'' stands for refactoring-oriented (and may be less useful for other analyses).

>>>>

\chapter{Defining refactoring-oriented clone types}\label{chap:clonetypes}
In Section~\ref{chap:backgroundclonetypes} we introduced the four clone types as defined in literature. These simple % TODO: Sander - why are they simple, might just remove that word
definitions are suitable for analysis of a codebase. Their detection results in simple to understand numbers to argue about a codebase. % TODO: Sander -this seems like an opinion
However, these clone types have a few flaws which makes it hard to argue to what extend two fragments of code can and should be refactored. For each of type 1-3 clones~\cite{roy2007survey} we list our solutions to their shortcomings to increase the chance that we can refactor the clone while improving the design.

We also look into clone detection tools for their suitability to support the proposed clone type definitions. We selected a few criteria % TODO: Sander -...?
Most clone detection tools support these definitions of clone types. However, many of these tools use a vastly different approach. A study by Saini et al \cite{saini2018towards} outlines different clone detection tools and compares their results for each of type 1-3 clones. Even though they operate on the same type definitions, the tools used in this study yield different results. % TODO: Sander - hier mist nog conclusie van dit stukje

\section{Shortcomings of clone types}
Clone types 1-3 (further explained in Section~\ref{chap:backgroundclonetypes}) allow reasoning about the duplication in a software system. Clones by these definitions can relatively easily and efficiently be detected. % TODO: Sander - compared to?
This has allowed for large scale analyses of duplication~\cite{livieri2007very}. However, these clone type definitions have shortcomings which makes the clones detected in correspondence with these definitions less valuable for (automated) refactoring purposes.

@@

\section{Refactoring-oriented clone types}\label{sec:rtypes}
To resolve the shortcomings of clone types as outlined in the previous section, we propose alternative definitions for clone types directed at detecting clones that can and should be refactored. We have named these clones T1R (type 1R), T2R and T3R clones. These definitions address problems of the corresponding literature definitions. The ``R'' stands for refactoring-oriented (and may be less useful for other analyses) % TODO: Sander - zou dat laatste remark miss weglaten of echt uitleggen.
